---
title: "linear model from Web"
author: "mariahlynn woolfolk"
date: "summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Intoduction   

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

### Methods  

all variables are numerical, so we use num~num helper file


## Model 1 SIM~ARMS

### Descriptive Results - Scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

This scatter plot shows that SIMS are higher when ARM strength increases.    

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation shows that 0.6860 is positive. Which means that higher strength would make the SIMS higher. Therefore, if the correlation were to be negative it would most likely decrease. 


### Inferential  (Build model.1)


```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

#### scatterplot with model fit

```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col='red')
```

The P-value for the intercept and slope is 0.00000000000000022. Therefore we reject the null hypothesis because it is less than 0.05    

### Predict at target point

```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.1, newdata, interval='pred')
```
The SIMS prediction for ARM=88 is 0.71, the predicting interval is (-1.73, 3.14)

## Model 2 SIM~GRIP

### Descriptive results -- Scatterplot 

```{r}
plot(SIMS~GRIP,data=data)
```

The scatter plot above shows that a higher GRIP strength also makes the simulation increase.    


### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```
This shows that when GRIP strength increases the simulation will become positive.

### Inferential  (Build model.2)
 
```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```

#### scatterplot with model fit

```{r}
plot(SIMS~GRIP,data=data)
abline(model.2, col='red')
```


The P-value for the interceptor and slope are 0.00000000000000022  



#### predict model.2 at target point


```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.2, newdata, interval='pred')
```

The SIMS prediction for GRIP=94 is 0.54, the predicting interval is (-3.10, 2.03)


## Model 3 SIM~ARM+GRIP

### Descriptive Results - Scatterplot

```{r}
plot(SIMS~ARM+GRIP, data=data)
```

### Numerical results

```{r}
df= data.frame(SIMS=data$SIMS,Grip=data$GRIP, ARM=data$ARM)
cor(df)
```

The comparison matrix shows that SIMS correlates with GRIP and ARM: And GRIP correlates with SIMS


### Numerical  reults (cor)

```{r}
cor(SIMS~ARM+GRIP, data=data)
```

This shows the similarity between the SIMS score and the variables ARM and GRIP.


### Inferential  (Build 2-dimentional model.3)

```{r}
model.3 = lm(SIMS ~ ARM + GRIP, data=data) 
summary.lm(model.3)
```

The p-value from the F-statistics is 0.0000000000000002, this means the data fits the model. Intercept p-value is 0.0000000000000002  
Arm p-value is 0.00000001  
Grip p-value is 0.00004  
Looking at this information they are all smaller than 0.05  





#### predict model.3 at target point

```{r}
newdata = data.frame(GRIP=94, ARM=88)
predSIMS = predict(model.3, newdata, interval='pred')
print(predSIMS)
```
The predicted SIMS score for GRIP=94, ARM=88 is 0.150

The interval prediction are (-2.13, 2.43)


```{r}
anova(model.1, model.3)
summary.lm(model.1)
model.1=lm(SIMS~ ARM, data=data)

```

The anova shows that there is a difference on the models since the p-value is 0.000004. 

The RSS in model 2 is 188 which is smaller than model 1 which is 217. 

This means that SIMS~ARM+GRIP is the better model because model 1 is SIMS ~ ARM only. 


### Model.2 vs Model.3

```{r}
model.2=lm(SIMS ~ GRIP, data=data)
anova(model.2, model.3)
summary.lm(model.1)
```

Anova shows that the Model SIM~ARM+GRIP is far better than the Model SIMS~GRIP. This is because the RSS is lower (188) and the p-value is 0.000000014

## Informally compare Model.1 with model.2

```{r}
summary.lm(model.1)
summary.lm(model.2)
```

Model.1 SIMS~ARM is better than model.2 SIMS~GRIP because the residual for model.1 was lower than model.2
