---
title: "linear model from Web"
author: "mariahlynn woolfolk"
date: "summer 2024"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Intoduction   

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

### Methods  

all variables are numerical, so we use num~num helper file


## Model 1 SIM~ARMS

### Descriptive Results - Scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

This scatter plot shows that SIMS are higher when ARM strength increases.    

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation shows that 0.6860 is positive. Which means that higher strength would make the SIMS higher. Therefore, if the correlation were to be negative it would most likely decrease. 


### Inferential  (Build model.1)


```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

#### scatterplot with model fit

```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col='red')
```

The P-value for the intercept and slope is 0.00000000000000022. Therefore we reject the null hypothesis because it is less than 0.05    

### Predict at target point

```{r}
newdata = data.frame(ARM =88, GRIP=94)
predict(model.1, newdata, interval='pred')
```
The SIMS prediction for ARM=88 is 0.71, the predicting interval is (-1.73, 3.14)

## Model 2 SIM~GRIP

### Descriptive results -- Scatterplot 

```{r}
plot(SIMS~GRIP,data=data)
```
The scatter plot above shows that a higher GRIP strength also makes the simulation increase.


### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```
This shows that when GRIP strength increases the simulation will become positive.

### Inferential  (Build model.2)
  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

